QPLIB: A Library of Quadratic Programming Instances

### 2017-02-07 QPLIB-3.tex
corrected tiny typo
mention that there is now also ampl .mod available

To evaluate the instances we decided, for practical reasons, to use
GAMS as common platform for all the experiments involving commercial and
non-commercial solvers. For this reason, we decided to translate all
instances into the GAMS format (\texttt{.gms}).
% In a preliminary phase, all the instances received were divided
%according to their format and subsequently translated.
%In \S\ref{subsec:tools}
%%~\ref{subsec:CP_convex}
%the tools used
%to translate an instance from a given format to the \texttt{.gms} format
%%to translate an instance from and to a given format to the
%\texttt{.gms} format
%are described in more detail.\\
%
In addition, we have introduced a new, specific \texttt{.qplib} format. This
new format is capable of describing all the instances of the library in a sparse form.
In comparison to a more \emph{high level} format like \texttt{.gms}, the new
format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed. See the \S\ref{sec:format} for more details.

@@

Similarly, Figures~\ref{fig:pic_constr} (left)
%Figure~\ref{fig:pic_constr_medium}
%and Figure~\ref{fig:pic_constr_big}
describes how the number of linear and quadratic
constraints are distributed within the instances. The instances are sorted
accordingly to the total number of constraints.
% and divided in the same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
For each instance we report the total number of constraints with a ``$+$''
and the total number of quadratic constraints
with a ``$\times$''. Also, in this case, different percentages of linear and
quadratic constraints are present and well-distributed across the spectrum
of constraint sizes, although both medium- and large-size instances show a
prevalence of lower percentages of quadratic constraints. In particular,
from Figure~\ref{fig:pic_constr} (left) we learn that while the maximum number
of linear constraints exceeds 100000, the maximum number of quadratic
constraints tops up at around 10000. This is, however, reasonable, considering
how quadratic constraints can, in general, be expected to be much more
computationally challenging than linear ones, especially if nonconvex.

@@

The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}, which was created by extending
scripts and tools initially developed for MINLPLib\,2~\cite{Vigerske2014}.
%
Beyond all instances in GAMS, \texttt{.lp}, and \texttt{.qplib} format, the website provides a rich set of metadata for each
instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of
different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
problem.
%
In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
Lagrangian function.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.

>>>>

To evaluate the instances we decided, for practical reasons, to use
GAMS as common platform for all the experiments involving commercial and
non-commercial solvers. For this reason, we decided to translate all
instances into the GAMS format (\texttt{.gms}).
% In a preliminary phase, all the instances received were divided
%according to their format and subsequently translated.
%In \S\ref{subsec:tools}
%%~\ref{subsec:CP_convex}
%the tools used
%to translate an instance from a given format to the \texttt{.gms} format
%%to translate an instance from and to a given format to the
%\texttt{.gms} format
%are described in more detail.\\
%
In addition, we have introduced a new, specific \texttt{.qplib} format. This
new format is capable of describing all the instances of the library in a sparse form.
In comparison to a more \emph{high level} and human readable format like \texttt{.gms}, the new
format presents three advantages: it is easier to read by a self-made parser, it produces smaller files, and the order of variables and constraints is fixed. See the \S\ref{sec:format} for more details.

@@

Similarly, Figure~\ref{fig:pic_constr} (left)
%Figure~\ref{fig:pic_constr_medium}
%and Figure~\ref{fig:pic_constr_big}
describes how the number of linear and quadratic
constraints are distributed within the instances. The instances are sorted
accordingly to the total number of constraints.
% and divided in the same manner as for Figures~\ref{fig:pic_var_small}--\ref{fig:pic_var_large}.
For each instance we report the total number of constraints with a ``$+$''
and the total number of quadratic constraints
with a ``$\times$''. Also, in this case, different percentages of linear and
quadratic constraints are present and well-distributed across the spectrum
of constraint sizes, although both medium- and large-size instances show a
prevalence of lower percentages of quadratic constraints. In particular,
from Figure~\ref{fig:pic_constr} (left) we learn that while the maximum number
of linear constraints exceeds 100000, the maximum number of quadratic
constraints tops up at around 10000. This is, however, reasonable, considering
how quadratic constraints can, in general, be expected to be much more
computationally challenging than linear ones, especially if nonconvex.

@@

The instances of QPLIB are publicly accessible at the website \url{http://qplib.zib.de}, which was created by extending
scripts and tools initially developed for MINLPLib\,2~\cite{Vigerske2014}.
%
Beyond all instances in GAMS, \texttt{.lp}, AMPL (\texttt{.mod}), and \texttt{.qplib} format, the website provides a rich set of metadata for each
instance: the three letter problem classification (as described in Section \ref{subsec:final set}), basic properties such as the number of variables and constraints of
different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
problem.
%
In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
Lagrangian function.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.
